---
title: "Example_usage"
output: html_document
date: '2023-11-20'
---

```{r setup}
# devtools::install()
library(MethQuant)
library(GenomicRanges)
```

Overview of the scores
```{r}
s1 <- getScores(level="sc", axis="w")
s1
```


# Single Cell 

dummy single cell data
```{r, dummy single-cell data}
md <- replicate(100, sample(c(0,1,NA), 1e4, replace=TRUE))
md <- as.data.table(md)
md$chr <- "1"
pos <- sample(1:1e6, 1e4)
pos <- pos[order(pos)]
md$pos <- pos
regions <- GenomicRanges::GRanges(seqnames=c("1"), ranges=IRanges(start=1, end=1e6))
```

Get plain score. All the score parameters go into the ... arguments (cdot). See getScores() for an overview of the parameters of the respective score
```{r, sc: plain score}
s <- methScores(md, score="trProb", axis="w", posCol="pos", seqCol="chr", 
                binMode=c("custom"), regions=regions,
                weighDist=FALSE)
head(s, 5)
```

Get score weighted by distance of CpGs
```{r, sc: weighted by distance}
s <- methScores(md, score="trProb", axis="w", posCol="pos", seqCol="chr", 
                binMode=c("custom"), regions=regions,
                weighDist=TRUE, decay="linear", maxDist=500)
head(s, 5)
```

Get score weighted by distance of CpGs & percentile within background distribution
```{r, sc: weighted by distance plus background}
s <- methScores(md, score="trProb", axis="w", posCol="pos", seqCol="chr", 
                binMode=c("custom"), regions=regions,
                weighDist=TRUE, decay="linear", maxDist=500, getBg=TRUE, 
                nIterations=2)
head(s, 5)
```
Calculate multiple scores and their percentiles
```{r, sc: several scores}
s <- methScores(md, score=c("trProb", "smc", "metLevel"), axis="w", posCol="pos", seqCol="chr", 
                binMode=c("custom"), regions=regions,
                weighDist=TRUE, decay="linear", maxDist=500, getBg=TRUE, 
                nIterations=2)
head(s, 5)
```

On height axis:
```{r, sc:height}
s <- methScores(md, score=c("metLevel", "shannonEn"), axis="h", posCol="pos", seqCol="chr", 
                binMode=c("custom"), regions=regions)
head(s, 5)
```

# Bulk 

```{r, dummy bulk data}
md <- replicate(100, runif(1e4))
md <- as.data.table(md)
md$chr <- "1"
pos <- sample(1:1e6, 1e4)
pos <- pos[order(pos)]
md$pos <- pos
regions <- GenomicRanges::GRanges(seqnames=c("1"), ranges=IRanges(start=1, end=1e6))
```

plain score
```{r, bulk: single score}
s <- methScores(md, score="sampEn", axis="w", posCol="pos", seqCol="chr", 
                binMode=c("custom"), regions=regions,
                weighDist=FALSE, naMode="keep", r=0.5, measure="euclidean")
head(s, 5)
```
change parametrization
```{r, bulk: single score, with different parameters}
s <- methScores(md, score="sampEn", axis="w", posCol="pos", seqCol="chr", 
                binMode=c("custom"), regions=regions,
                weighDist=FALSE, naMode="keep", r=2, measure="euclidean")
head(s, 5)
```

weighted by distance (slow for sampEn, especially in the bulk case+weighted, not recommended to run with backgrounds)
```{r, bulk: distance weighted score}
s <- methScores(md, score="corr", axis="w", posCol="pos", seqCol="chr", 
                binMode=c("custom"), regions=regions,
                weighDist=TRUE)
head(s, 5)
```

with background 
```{r, bulk: distance weighted score with background}
s <- methScores(md, score="corr", axis="w", posCol="pos", seqCol="chr", 
                binMode=c("custom"), regions=regions,
                weighDist=TRUE, 
                getBg=TRUE, nIterations=2)
head(s, 5)
```

several scores
```{r, bulk: several scores}
s <- methScores(md, score=c("sampEn", "corr", "metLevel"), axis="w", posCol="pos", seqCol="chr", 
                binMode=c("custom"), regions=regions,
                weighDist=FALSE, naMode="keep", r=0.5, measure="euclidean")
head(s, 5)
```

height axis
```{r}
s <- methScores(md, score=c("metLevel", "shannonEn"), 
                axis="h", posCol="pos", seqCol="chr", 
                binMode=c("custom"), regions=regions)
head(s, 5)
```